### 2022-06-27 draft.tex
Deploying to gh-pages from @ PoisotLab/ms_betacov_mapping@8d14abcf17db4e1500203637797d46cbe3a3af37 🚀

On closer inspection, these patterns recapitulate the evolutionary
history of both the order Chiroptera and the genus
\emph{Betacoronavirus}. Horseshoe bats (Rhinolophidae) are both the
reservoirs of the SARS-like viruses (subgenus \emph{Sarbecovirus}), the
group of pandemic threats that have been of the greatest interest to
researchers\textsuperscript{17} (and so have been sampled most
intensively).\textsuperscript{16} The hotspots of host richness and
viral diversity in southeast Asia---both of which are disproportionately
high, considering the global landscape of bat species richness---are
almost entirely driven by viral adaptive radiation through host
switching within this clade\textsuperscript{3,14}. In contrast, the
Neotropical hotspot of viral distinctiveness is driven by isolation by
host vicariance. Out of the four main groups of betacoronaviruses, only
the subgenus \emph{Merbecovirus} (MERS-like viruses) has been found in
animals in the Americas---an introduction that is generally presumed to
be ancient.\textsuperscript{3} While comparatively understudied, New
World merbecoviruses have been found in the ghost-faced bats
(Mormoopidae), New World leaf-nosed bats (Phyllostomidae), and
free-tailed bats (Molossidae) (add cite: Olival 2020 PLoS Pathogens).
The former two groups are endemic to the Neotropics, while the explosive
adaptive radiations of the latter two (and particularly the
phyllostomids) are responsible for the hotspot of bat diversity in the
Amazon. Together, these clades of New World bats play host to a distinct
regime of betacoronavirus coevolution.

@@

Overall, these results suggest that, although there are unique hotspots
of viral adaptive radiation, biogeographic regions are mostly consistent
between bats and their viruses. This may be surprising, given that
cospeciation plays a minor role in coronavirus
diversification,\textsuperscript{2} which could theoretically allow for
broad divergence in their biogeography. However, host jumps at the
family level or higher are relatively rare and significant events in
coronavirus evolutionary history;\textsuperscript{2,17} as a result, we
suggest the mosaic of betacoronavirus phylogeography is assembled from a
set of overlapping smaller coevolutionary systems, superimposed in space
and filtered by the importance of different subgroups in local host
communities. For example, the most speciose and cosmopolitan family of
bats, the vesper bats (Vespertilionidae), are considered the primary
reservoir of merbecoviruses;\textsuperscript{3,17} but in the Americas,
where merbecoviruses are the only lineage present, they have only been
found in other bat taxa. At the coarsest scale, these heterogeneities
are lost, and betacoronavirus biogeography tracks the deep rifts in bat
evolutionary history---but within these broad regions, the component
coevolutionary systems can have very different dynamics.

>>>>

Overall, these patterns recapitulate the evolutionary history of both
the order Chiroptera and the genus \emph{Betacoronavirus}. Horseshoe
bats (Rhinolophidae) are both the reservoirs of the SARS-like viruses
(subgenus \emph{Sarbecovirus}), the group of pandemic threats that have
been of the greatest interest to researchers\textsuperscript{17} (and so
have been sampled most intensively).\textsuperscript{16} The hotspots of
host richness and viral diversity in southeast Asia---both of which are
disproportionately high, considering the global landscape of bat species
richness---are almost entirely driven by viral adaptive radiation
through host switching within this clade\textsuperscript{3,14}. In
contrast, the Neotropical hotspot of viral distinctiveness is driven by
isolation by host vicariance. Out of the four main groups of
betacoronaviruses, only the subgenus \emph{Merbecovirus} (MERS-like
viruses) has been found in animals in the Americas---an introduction
that is generally presumed to be ancient.\textsuperscript{3} While
comparatively understudied, New World merbecoviruses have been found in
the ghost-faced bats (Mormoopidae), New World leaf-nosed bats
(Phyllostomidae), and free-tailed bats (Molossidae) (add cite: Olival
2020 PLoS Pathogens). The former two groups are endemic to the
Neotropics, while the explosive adaptive radiations of the latter two
(and particularly the phyllostomids) are responsible for the hotspot of
bat diversity in the Amazon. Together, these clades of New World bats
play host to a distinct regime of betacoronavirus coevolution.

@@

Overall, these results suggest that, although there are unique hotspots
of viral adaptive radiation, biogeographic regions are mostly consistent
between bats and their viruses. This may be surprising, given that
cospeciation plays a minor role in coronavirus
diversification,\textsuperscript{2} which could theoretically allow for
broad divergence in their biogeography. However, host jumps at the
family level or higher are relatively rare and significant events in
coronavirus evolutionary history;\textsuperscript{2,17} as a result, the
mosaic of betacoronavirus phylogeography is assembled from a set of
overlapping smaller coevolutionary systems, superimposed in space and
filtered by the importance of different subgroups in local host
communities. For example, the most speciose and cosmopolitan family of
bats, the vesper bats (Vespertilionidae), are considered the primary
reservoir of merbecoviruses;\textsuperscript{3,17} but in the Americas,
where merbecoviruses are the only lineage present, they have only been
found in other bat taxa. At the coarsest scale, these heterogeneities
are lost, and betacoronavirus biogeography tracks the deep rifts in bat
evolutionary history---but within these broad regions, the component
coevolutionary systems can have very different dynamics.
